---
title: "Graphing the data"
author: "Asmus Hylkjaer"
date: " created 07/12/2022, updated date: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
---
##Graphing
dunno what this does
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE)
```

```{r library}
library(tidyverse)
library(readxl)
library(here)
```


```{r load-data}
Heritage_list_graph <- read_excel("data/whc-sites-2021.xls")
Heritage_list_graph_data <-read_delim("data/inscribtion_data.csv")
Heritage_list_graph_data2 <- read_excel("data/inscribtion_data.xlsx")

```

## Inspect the data
```{r inspect}
head(Heritage_list_graph)
head(Heritage_list_graph_data)
head(Heritage_list_graph_data2)
```
## Let's how many percentage are added a year pr. region
```{r plot-yearly-percentage, eval = TRUE}
ggplot(Heritage_list_graph_data2, aes(x=year)) + 
  geom_line(aes(y=procentage_year_Eu, color="EU & NA")) + 
  geom_line(aes(y=procentage_year_AsP, color="Asia & pasific")) + 
  geom_line(aes(y=procentage_year_AF, color="Africa"))+
  geom_line(aes(y=procentage_year_AS, color="Arab states")) + 
  geom_line(aes(y=procentage_year_LA, color="Latin america"))+
  labs(title="World heritage sites added pr. year percentages pr. region",
       y="Percentage",x="Year",color="Region")+
  scale_x_continuous(breaks = seq(1978,2021,5))+
  theme(axis.text.x = element_text(face="bold", color="#993333", 
                                   size=8, angle=0),
        axis.text.y = element_text(face="bold", color="#993333", 
                                   size=8, angle=0))

```


## Let's how many percentag
```{r plot-cul-percentage, eval = TRUE}
ggplot(Heritage_list_graph_data2, aes(x=year)) + 
  geom_line(aes(y=Eu_percentage_cul, color="EU & NA")) + 
  geom_line(aes(y=AsP_percentage_cul, color="Asia & pasific")) + 
  geom_line(aes(y=AF_percentage_cul, color="Africa"))+
  geom_line(aes(y=AS_percentage_cul, color="Arab states")) + 
  geom_line(aes(y=LA_percentage_cul, color="Latin america"))+
  labs(title="World heritage sites regions split into percentages",
       y="Percentage",x="Year",color="Region")+
  scale_x_continuous(breaks = seq(1978,2021,5))+
  theme(axis.text.x = element_text(face="bold", color="#993333", 
                                   size=8, angle=0),
        axis.text.y = element_text(face="bold", color="#993333", 
                                   size=8, angle=0))

```


notes for the excel sheet
wanted to do csv. didnt work
did manual editing of the spreadshet. but was using ecxel funktions. no numbers were added manually.
added collum and renamed them to make them easier to work with.
so the data is imported from unesco 

(https://whc.unesco.org/en/list/stat):
"Number of World Heritage properties inscribed each year by region"

added collums
anything mentioning cul(minative) or percentage 



gdp - cost money to get a site
look into countries asking to get sites added to the list 
 - might have other priorities